---
title: "Data Visualisation - Analsying Lending Club's Loans"
author: "Prashorst Meas"
date: "January 22, 2019"
output: 
  html_document:
    toc: true
    number_sections: true
---
# Introduction
  
Data visualisation is an indispensible tool for data scientists to make sense of the underlying data and to reveal hidden insights and trends through *Exploratory Data Analysis* (EDA). It is also used to confirm the findings from machine learning models to corroborate results and establish a conclusion.

## Data Source

The data set used in this data visualisation exercise is derived from [Kaggle](https://www.kaggle.com/wendykan/lending-club-loan-data) which contained **Lending Club Loan Data** from 2007 to 2015.

## Format

This report is written using **R Markdown** as a formatting syntax to create an **HTML** document through `Knitr` and `Pandoc`. The dynamic report features static graphs produced from `ggplot2` and interactive data visualisation generated by `Shiny`.   

## Exploratory Data Analysis

To begin the data visualisation project, the data set should be initially loaded into the R session. Since this dataset is already available on the web, it can be directly imported using a function to read the **csv** file.

Here is where the first problem manifests. Deciding on what function from which package to use depend on several factors such as the speed of parsing, the data type rendered, and the default specifications.

The 3 most popular functions to import data are `read.csv(file.extension)` in base R from the `utils` package, using `read_csv(file.extension)` from `readr` package, or `fread(file.extension)` from `data.table` package.

For this project, `fread` is the most suitable for the following reasons:

* It is proven to have the fastest when parsing data, especially for large files
* It can be configured to be parsed as a data frame by setting the argument `data.table = FALSE` which is useful for data manipulation using `dplyr` and data visualisation using `ggplot2`

The first step is setting up the working directory and loading the nescessary packages.

```{r message = FALSE}
setwd("/Users/apple/Desktop/Data Science Projects/lending-club-loan-data")

library(data.table) 
library(dplyr) # data manipulation
```

The data set is then imported and the class of the result is checked to ensure the code is working as intended.

```{r}
loan_data <- fread("/Users/apple/Desktop/Data Science Projects/lending-club-loan-data/loan.csv", data.table = FALSE)

class(loan_data)
```

To get a sense of the size of the data, let's start exploring and diving into its content.

```{r}
dim(loan_data)

names(loan_data)
```

Next, different kinds of summaries can be performed on the data.

```{r results = "hide"}
glimpse(loan_data)

summary(loan_data)
```

